Stringy canonical forms and binary geometries from associahedra, cyclohedra and generalized permutohedra

### 2020-07-16 draft_ABP.tex
Updates from Overleaf

The polytope ${\cal P}$ is bounded by inequalities $W_a^JS_J\geq 0$ in the space of $\mathbf X$, where $W_a^J$ is determined by the facet $F_a$ of $\mathcal P$.  These inequalities also cut out a polyhedron in the $(d+m)$-dimensional space of $\mathbf S$, which is defined to be the big polyhedron of $\mathcal P$, 
and the integral $\mathcal I(\mathbf S)$ converges for each point $\mathbf S$ 
inside this polyhedron. Dually, we can describe the big polyhedron by its vertices. Suppose 
$\{V_J^\alpha\}_{1\leq \alpha \leq v}$ are its non-zero vertices ($v$ denotes the number of vertices), a point $\mathbf S$ inside 
the polyhedron can be written as $S_J = F_\alpha V^\alpha_J$ with non-negative coefficients 
$\{F_\alpha\}_{1\leq \alpha \leq v}$.  Therefore, the integral $\mathcal I(\mathbf S)$
can be seen as a function of ``dual coordinates'' $\{F_\alpha\}$ which encourages us to 
rewrite the integral to make it manifest by (here we collectively denote $p_J=(x_1, \cdots, x_d, p_1, \cdots, p_m)$)~\footnote{Note that one can indeed treat the monomials $x_i^{\alpha' X_i}$ on the equal footing as polynomials, since Minkowski sum w.r.t. them amounts to merely translating the polytope $c_I N[p_I]$ by $-X_i$ in the $i$-th direction~\cite{Arkani-Hamed:2019mrd}.}
\begin{equation}
	\mathcal I(\mathbf F)=
	\int_{\mathbb R_+^d} \prod_{i=1}^d \frac{d x_i}{x_i}~%\frac{\mathrm d x_1\cdots \mathrm d x_n}{x_1\cdots x_n}
	\prod_{\alpha}u_\alpha^{\alpha' F_\alpha}
	=
	\int_{\mathbb R_+^d}\prod_{i=1}^d \frac{d x_i}{x_i}~%\frac{\mathrm d x_1\cdots \mathrm d x_n}{x_1\cdots x_n}
	\prod_{J}^{d{+}m} p_J^{\alpha' S_J},
\end{equation}
where we introduce a new set of variables $u_\alpha = \prod_J p_J^{V_J^\alpha}$ for 
$\alpha=1,\dots, v$. Note that there's no unique way to write $S_J=F_\alpha V_J^\alpha$ if the big polyhedron is not a simplex, \textit{i.e.} $v>d+m$, and in this case,
$u$ variables $\{u_\alpha \}$ are not multiplicatively independent. Therefore, it is an important special case when the big polyhedron is a simplex, \textit{i.e.} $N=v=d+m$, where $N$ is 
the number of facets of the big polyhedron or the original polytope $\mathcal P$,
and inequalities $F_\alpha=S_J(V^{-1})^J_\alpha \geq 0$ are exactly the inequalities $W_a^J S_J\geq 0$ up to resales $F_\alpha \mapsto t F_\alpha$ for positive factors $t$, because
there's no nontrivial linear isometry from the simplex to itself besides permutations of vertices. Therefore, we identity the label of facets $\alpha$ with the label of 
vertices $J$ and get $V=W^{-1}$. Furthermore, every facet of the original polytope $\mathcal P$ can be associated with a single $u_\alpha$ going to zero. 
%Generally, the big polyhedron of a stringy integral is not necessary to be a simplex.

@@

These nice properties become manifest in the configuration space: not only the big polyhedron is a simplex, the $u$ equations are of very special form:
\begin{equation}\label{perfectu}
	1-u_\alpha=\prod_{\beta} u_\beta^{\beta||\alpha},
\end{equation}
where the product is over all $N$ variables ($N$ facets), and the integer $\beta || \alpha\geq 0$ is the so-called compatibility degree defined in the corresponding cluster algebra~\cite{Zelevinskysys}. Varieties defined by $u$ equations associated with finite type clusters  are called cluster configuration spaces.
We see that when $u_\alpha$ (for a facet $\alpha$) goes to $0$, it forces all the $u_\beta$ with $\beta || \alpha>0$ goes to $1$ (for facets $\beta$ are exactly those {\it incompatible} with $\alpha$); the remaining $u$'s that are compatible with $u_\alpha$ satisfy $u$ equations for the facet, which factorizes as removing a node of the Dynkin diagram. Therefore, we say that the configuration space is a {\it binary geometry}. Note that this is true also for the complex case, and if we further require all $u$ variables to be positive, we have $0<u<1$ which gives the positive configuration space~\cite{Arkani-Hamed:2020cig}. The binary property is what guarantees the stringy integral to factorize even at finite $\alpha'$: on the pole $X_\alpha \to 0$, $u_\alpha\to 0$ forces all incompatible $u_\beta\to 1$ which decoupled from $\prod u^{\alpha' X}$, and the integral becomes the factorized one for the facet $\alpha$.

@@

In this paper we introduce a large class of integrals,  ${\cal I}_{\bf B}$, which can be regarded as a family of rigid stringy canonical forms%~\cite{Arkani-Hamed:2019mrd} 
for generalized permutohedra ${\mathscr P}(\bf B)$, and we show that corresponding configuration spaces are always binary geometries. These rigid integrals satisfy the remarkable property that, at any ``massless'' pole, {\it i.e.} poles of the leading order $\Omega({\mathscr P}(\bf B) )$, the residue is given by a product of two such integrals, ${\cal I}_{{\bf B}_L} \times {\cal I}_{{\bf B}_R}$. Our results greatly extend the scope of binary geometries and associated integrals with such ``perfect'' factorization properties, which were originally thought to tie to the more special class of generalized associahedra for finite-type cluster algebras. For infinite families of cases as degenerations of associahedra and cyclohedra (which belong to both generalized permutohedra and generalized associahedra), we find additionally that the $u$ equations are ``perfect'' just like the case for cluster configuration spaces. Our preliminary investigations have opened up various new avenues to be explored, and we end with listing some open questions.

>>>>

The polytope ${\cal P}$ is bounded by inequalities $W_a^JS_J\geq 0$ in the space of $\mathbf X$, where $W_a^J$ is determined by the facet $F_a$ of $\mathcal P$.  These inequalities also cut out a polyhedron in the $(d+m)$-dimensional space of $\mathbf S$, which is defined to be the big polyhedron of $\mathcal P$, 
and the integral $\mathcal I(\mathbf S)$ converges for each point $\mathbf S$ 
inside this polyhedron. Dually, we can describe the big polyhedron by its vertices. Suppose 
$\{V_J^\alpha\}_{1\leq \alpha \leq v}$ are its non-zero vertices ($v$ denotes the number of vertices), a point $\mathbf S$ inside 
the polyhedron can be written as $S_J = F_\alpha V^\alpha_J$ with non-negative coefficients 
$\{F_\alpha\}_{1\leq \alpha \leq v}$.  Therefore, the integral $\mathcal I(\mathbf S)$
can be seen as a function of ``dual coordinates'' $\{F_\alpha\}$ which encourages us to 
rewrite the integral to make it manifest by (here we collectively denote $p_J=(x_1, \cdots, x_d, p_1, \cdots, p_m)$)~\footnote{Note that one can indeed treat the monomials $x_i^{\alpha' X_i}$ on the equal footing as polynomials, since Minkowski sum w.r.t. them amounts to merely translating the polytope $c_I N[p_I]$ by $-X_i$ in the $i$-th direction~\cite{Arkani-Hamed:2019mrd}.}
\begin{equation}
	\mathcal I(\mathbf F)=
	\int_{\mathbb R_+^d} \prod_{i=1}^d \frac{d x_i}{x_i}~%\frac{\mathrm d x_1\cdots \mathrm d x_n}{x_1\cdots x_n}
	\prod_{\alpha}u_\alpha^{\alpha' F_\alpha}
	=
	\int_{\mathbb R_+^d}\prod_{i=1}^d \frac{d x_i}{x_i}~%\frac{\mathrm d x_1\cdots \mathrm d x_n}{x_1\cdots x_n}
	\prod_{J}^{d{+}m} p_J^{\alpha' S_J},
\end{equation}
where we introduce a new set of variables $u_\alpha = \prod_J p_J^{V_J^\alpha}$ for 
$\alpha=1,\dots, v$. Note that there's no unique way to write $S_J=F_\alpha V_J^\alpha$ if the big polyhedron is not a simplex, \textit{i.e.} $v>d+m$, and in this case,
$u$ variables $\{u_\alpha \}$ are not multiplicatively independent. Therefore, it is an important special case when the big polyhedron is a simplex, \textit{i.e.} $N=v=d+m$, where $N$ is 
the number of facets of the big polyhedron or the original polytope $\mathcal P$,
and inequalities $F_\alpha=S_J(V^{-1})^J_\alpha \geq 0$ are exactly the inequalities $W_a^J S_J\geq 0$ up to rescaling $F_\alpha \mapsto t F_\alpha$ for positive factors $t$, because
there's no nontrivial linear isometry from the simplex to itself besides permutations of vertices. Therefore, we identity the label of facets $\alpha$ with the label of 
vertices $J$ and get $V=W^{-1}$. Furthermore, every facet of the original polytope $\mathcal P$ can be associated with a single $u_\alpha$ going to zero. 
%Generally, the big polyhedron of a stringy integral is not necessary to be a simplex.

@@

These nice properties become manifest in the configuration space: not only the big polyhedron is a simplex, the $u$ equations are of very special form:
\begin{equation}\label{perfectu}
	1-u_\alpha=\prod_{\beta} u_\beta^{\beta||\alpha},
\end{equation}
where the product is over all $N$ variables ($N$ facets), and the integer $\beta || \alpha\geq 0$ is the so-called compatibility degree defined in the corresponding cluster algebra~\cite{Zelevinskysys}. Varieties defined by $u$ equations associated with finite type clusters  are called cluster configuration spaces.
We see that when $u_\alpha$ (for a facet $\alpha$) goes to $0$, it forces all the $u_\beta$ with $\beta || \alpha>0$ to go to $1$ (for facets $\beta$ are exactly those {\it incompatible} with $\alpha$); the remaining $u$'s that are compatible with $u_\alpha$ satisfy $u$ equations for the facet, which factorizes as removing a node of the Dynkin diagram. Therefore, we say that the configuration space is a {\it binary geometry}. Note that this is true also for the complex case, and if we further require all $u$ variables to be positive, we have $0<u<1$ which gives the positive configuration space~\cite{Arkani-Hamed:2020cig}. The binary property is what guarantees the stringy integral to factorize even at finite $\alpha'$: on the pole $X_\alpha \to 0$, $u_\alpha\to 0$ forces all incompatible $u_\beta\to 1$ which decoupled from $\prod u^{\alpha' X}$, and the integral becomes the factorized one for the facet $\alpha$.

@@

In this paper we introduce a large class of integrals,  ${\cal I}_{\bf B}$, which can be regarded as a family of rigid stringy canonical forms %~\cite{Arkani-Hamed:2019mrd} 
for generalized permutohedra ${\mathscr P}(\bf B)$, and we show that corresponding configuration spaces are always binary geometries. These rigid integrals satisfy the remarkable property that, at any ``massless'' pole, {\it i.e.} poles of the leading order $\Omega({\mathscr P}(\bf B) )$, the residue is given by a product of two such integrals, ${\cal I}_{{\bf B}_L} \times {\cal I}_{{\bf B}_R}$. Our results greatly extend the scope of binary geometries and associated integrals with such ``perfect'' factorization properties, which were originally thought to tie to the more special class of generalized associahedra for finite-type cluster algebras. For infinite families of cases as degenerations of associahedra and cyclohedra (which belong to both generalized permutohedra and generalized associahedra), we find additionally that the $u$ equations are ``perfect'' just like the case for cluster configuration spaces. Our preliminary investigations have opened up various new avenues to be explored, and we end with listing some open questions.
